---
title: "Toward A Beta Model:  A Tutorial For Using Beta Regression In Pychological Research To Analyze Proportional and Percentage Data"
shorttitle: "Beta Regression Tutorial"
author:
  - name: Jason Geller
    corresponding: true
    orcid: 0000-0002-7459-4505
    email: drjasongeller@gmail.com
    affiliations:
      - name: Boston College
        department: Department of Psychology and Neuroscience
        address: McGuinn 300
        city: Chestnut Hill
        region: MA
        country: USA
        postal-code: 02467
  - name: Robert Kubinec
  - name: Matti Vuorre
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: "Rates, percentages, and proportional data are widespread in psychology. These data are usually analyzed with methods falling under the general linear model, which are not ideal for this type of data. A better alterantive is the beta regession model which is based on the beta distribution. A beta regression can be used to model data that is non-normal, heteroscedastic, and bounded between an interval [0,1]. Thus, the beta regression model is well-suited to examine outcomes in psycholgical research expressed as proportions, percentages, or ratios. The overall purpose of this tutorial is to give researchers a hands-on demonstration of how to use beta regression using a real example from the psychological literature. First, we introduce the beta distribution and the beta regression model highlighting crucial components and assumptions. Second, we highlight how to conduct a beta regression in R  using an example dataset from the learning and memory literature. Some extensions of the beta model are then  discussed (e.g., zero-inflated, zero- one-inflated, and ordered beta). We present accompanying R code throughout. All code to reproduce this paper can be found on Github: "
keywords: [Beta regression, tutorial, cognitive psychology, learning and memory]
bibliography: bibliography.bib
format:
  apaquarto-docx: default
  apaquarto-html: default
execute: 
  warning: false
  message: false
---

In psychological research it is quite common to use response outcomes that are rates, percentages, or proportional in nature. For example, it is quite common in educational and cognitive research to measure learning by looking at accuracy on a final test. To make this more concrete, let’s say you run a memory experiment wherein participants read a short passage pertaining to some topic or subject. After a short distractor task, participants completed a final memory test consisting of 10 short-answer questions, each worth a different number of points (e.g., question 1 may be worth 4 points while question 2 may be worth 1 point). Your estimand of interest is the total number of correct answers out of the total possible points for each question. With this as your main outcome, the resulting metric would be proportional in nature. How do you analyze this type of data?

Statistical analysis of proportions can present numerous difficulties. By definition, proportions are limited to numerical values between, and including, 0 and 1, and the variability in the observed proportions usually varies systematically with the mean of the response. It is quite common to analyze proportional outcomes using methods falling under the general linear model (GLM). There are several issues with this approach, however. First, the GLM assumes the residuals in the model are normal distributed. Second, it assumes an unbounded distribution that can extend from -$\infty$ to $\infty$. Lastly, the GLM assumes constant residuals across varying levels of variables in the model. These assumptions are often violated when dealing with proportional data, which are typically bounded between 0 and 1 and may not follow a normal distribution [@ferrari2004; @paolino2001]. Adopting a model that does not capture your data accurately can have deleterious consequences, such as missing a true effect when it exists (Type 2 error), or mistaking an effect as real when it is not (Type 1 error). A goal for any researcher trying to draw inferences from their data is to fit a model that accurately captures the important features of the data and has predictive utility [@yarkoni2017].

The issues related to analyzing proportional data are not new (see [@bartlett1936]). Luckily, several analysis strategies are available to deal with them. One such approach we highlight here is beta regression [@ferrari2004; @paolino2001] and some of its alternatives. With the combination of open-source programming languages like R [@R] and the great community of package developers, it is becoming trivial to run analyses like beta regression. However, adoption of these methods, especially in psychology, is sparse. A quick Web of Science search for a 10 year period spanning 2014-2024 using (TS=(Psychology)) AND TS=(beta regression) as search terms returned fewer than 20 articles. One reason for the lack of adaptation could be the lack of resources available to wider community (but see [@heiss2021]; [@vuorre2019]; [@bendixen2023]). We attempt to the rectify this herein.

In this article, we plan to (a) give a brief, non-technical overview of the principles underlying beta regression, (b) walk-through an empirical example of applying beta regression using popular frequentist and Bayesian packages in the popular R programming language and (c) highlight the the extensions which are most relevant to researchers in psychology (e.g., zero-inflated, zero-one-inflated, and ordered beta regressions).

To make this tutorial useful, we will use the popular open source programming language R [@R]. We integrate key code chunks into the main text throughout so the reader can follow along. This manuscript is fully reproducible as it is written in quarto and is publicly available:.

Because frequentist statistics are still quite popular in psychology we show to perform beta regression using frequentist packages like `betareg` [@betareg-4] and `glmmTMB` [@glmmTMB-2] (if you have nested type data or working with more complex models). We, however, also highlight how to run these models from a Bayesian framework using the `brms` [@brms-3] package. Our main goal for this tutorial is be maximally useful regardless of statistical proclivities.

## Beta distribution

Going back to our example from the introduction, our main measure is continuous varying between 0 and 1. Given this, what kind of distribution can be used to fit this data? The beta distribution is perfect for analyzing outcomes like proportions, percentages, and ratios.[^1] The beta distribution has some desirable characteristics that make it ideal for analyzing this type of data: It is continuous, it is limited to numbers that fall between 0 and 1, and it highly flexible—it can take on a number of different distribution shapes. It is important to note that the beta distribution *excludes* numbers that are exactly 0 and exactly 1. That is, it cannot model values that are exactly 0 or 1.

[^1]: Unlike some other popular distributions (Gaussian, poisson, binomial) the beta distribution is not generally thought of as part of the exponential/GLM family.

The beta distribution can take on a number of different shapes. The location, skew, and spread of the distribution is controlled by two parameters: *shape1* and *shape2*. `Shape 1` is sometimes called $\alpha$ and `shape 2` is sometimes called $\beta$. Together these two parameters shape the density curve of the distribution. For example, let's suppose a participant got 4 out of 6 correct on a test item. We can take the number of correct on that particular test item (4) and divide that by the number of correct (4) + number of incorrect (2) and plot the resulting density curve. Shape1 in this example would be 4 (number of points received). This parameter reflects the number of successes. `Shape2` would be 2--the number of points not received. This parameter reflects the number of failures. Looking at Figure 1a, we see the distribution for one of our questions is shifted towards one indicating higher accuracy on the exam. If we reversed the values of the two parameters, we would get a distribution shifted towards 0 (Figure 1b), indicating a lower accuracy. By adjusting the values of two parameters, we can get a wide range of distributions (e.g., u-shaped, inverted u-shaped , normal, or uniform). As can be seen, the beta distribution is a distribution of proportions or probabilities.

```{r}
#| label: Figure 1
#| fig-cap: "A. Beta distribution for test item with 4 correct and 2 incorrect. B. Beta distribution for test item with 2 correct and 4 incorrect"
#| echo: false
#| message: false

library(tidyverse)
library(easystats)
library(patchwork)
library(cowplot)

p1 <- ggplot() +
  geom_function(fun = dbeta, args = list(shape1 = 4 , shape2 = 2),
                aes(color = "Beta(shape1 = 4, shape2 = 2)")) +
  scale_color_viridis_d(option = "plasma", name = "") +
  theme_lucid(base_size=18) +
  theme(legend.position = "bottom")

p2 <- ggplot() +
  geom_function(fun = dbeta, args = list(shape1 = 2 , shape2 = 4),
                aes(color = "Beta(shape1 = 4, shape2 = 2)")) +
  scale_color_viridis_d(option = "plasma", name = "") +
  theme_lucid(base_size=18) +
  theme(legend.position = "bottom")

plot_grid(p1, p2, labels = c('A', 'B'))

```

However, the canonical parametrization of $\alpha$ and $\beta$ does not lend itself to the regression framework. Thus, it is quite common to talk about $\mu$ and $\phi$ instead, where $\mu$ represents the mean or average, and $\phi$ represents the precision. We can reparamterize $\alpha$ and $\beta$ into $\mu$ and $\phi$:

$$
\begin{equation}
\begin{aligned}[t]
\text{Shape 1:} && a &= \mu \phi \\
\text{Shape 2:} && b &= (1 - \mu) \phi
\end{aligned}
\qquad\qquad\qquad
\begin{aligned}[t]
\text{Mean:} && \mu &= \frac{a}{a + b} \\
\text{Precision:} && \phi &= a + b
\end{aligned}
\end{equation}
$$\

The variance is a function of $\mu$ and $\phi$:

$$
\frac{\mu \cdot (1 - \mu)}{1 + \phi}
$$

## Beta regression

Beta regression provides a means to relate predictors to the parameters of a beta-distributed response variable. The beta regression is a two-part model model that It utilizes a link function (commonly a logit) to model the mean of our response variable as a function of the predictor variables, while another link function models the dispersion parameter (commonly a log). This approach respects the bounded nature of the data and allows for heteroskedasticity, making it highly appropriate for data that represents proportions or rates.

# Example

## Data and Methods

Now we have built up an intuition about the beta distribution we can start to analyze some data. The principles of beta regression are best understood in the context of a real data set. The example we are gonna use comes from the learning and memory literature. A whole host of literature has shown extrinic cues like fluency (i.e., how easy something is to process) can influence metamemory (i.e., how well we think we will remember something). As an interesting example, when an instructor uses non-verbal gestures, has variable voice dynamics, moves about the space, and includes appropriate pauses when delivering content they are previeved as more fluent and this can influence our perceptions of how well we will remember something, but does not influence actual memory performance, or what we learn [@carpenter2013; @toftness2017; @witherby2022]. Seeing a fluent instructor makes it more likely individuals will *think* they will remember more content, but it has an actual weak effect on actual memory performance. However, in one study @wilford2020 showed fluency impacts actual memory. In several experiments, participants watched an instructor discussing the genetics of calico cats and mechanisms of skin wrinkling. The instructor was either fluent and disfluent. Participants who wacthed the fluent videos remembered more information on a final test compared to participants who watched a disfluent instructor discussing the same things. Given the interesting, and contradictory results, we chose this paper to highlight. In the current tutorial we are going to re-analyze the final recall data from Wilford et al. (2021; Experiment 1a). Data for the paper can be found here: <https://osf.io/6tyn4/>. Accuracy data is widely used in psychology and is well suited for beta regression. Despite this, it is common to treat accuracy data as continuous and unbounded, and analyze the resulting proportions using methods that fall under the general linear model. Below we will reproduce the analysis conducted by @wilford2020 (Experiment 1a) and then re-analyze it using beta regression. We hope to show how beta regression and its extensions can be a more powerful tool in making inferences about your data.

To recapitulate, Wilford et al. (2020; Expt 1a) presented participants with two short videos on the genetics of calico cats and why skin wrinkles. Participants viewed either disfluent or fluent versions of these videos. See examples here:. For each video, metamemory was assessed using judgement of learning (JOLS). JOLs require participants to rate an item on scale between 0-100 with 0 representing the item will not be remembered and a 100 representing they will definitely remember the item. In addition, other questions about the instructor were assessed and how much they learned. After a distractor task, a final free recall test was given were participants had to recall as much information about the video as they could in 3 minutes. Participants could score up to 10 points for each video. Here we will only being looking at the final recall data, but we could also analyze the JOL data as well with a beta rergession.

## Reanalysis of Wilford et al. Experiment 1a

### GLM approach

In Experiment 1a, @wilford2020 only used the first time point and compared fluent and disfluent conditions with a *t*-test. In our re-analysis, we will also run a *t*-test, but in a regression context. This allows for easier generalization to the beta regression approach. Specifically, we will examine accuracy on final test (because the score was on a 10 point scale we multiplied each value by 10 and divided by 100 to get a proportion) as our DV and Condition. Fluency will be dummy coded with the fluent level as our reference level.

### Load packages and data

As a first step, we will load the necessary packages we will be using and the data. While we load all the necessary packages here, I highlight when packages are needed when the analyses are run.

```{r}
#| echo: true
#| message: false
#| warning: false


# packages needed
library(tidyverse) # tidy functions/data wrangling/viz
library(betareg) # run beta regression 
library(glmmTMB) # zero inflated beta
library(easystats)
library(gghalves)
library(ggbeeswarm)       # Special distribution-shaped point jittering
library(scales) # percentage
library(tinytable) # tables
library(marginaleffects) # marginal effects
library(extraDistr)       # Use extra distributions like dprop()
library(brms) # bayesian models



options(scipen = 999) # get rid of scienitifc notation
```

Next, we load in our data, rename the columns to make them more informative, and transform the data. Here we transform accuracy so it is a proportion by multiplying each score by 100. FInally we dummy code the fluency variable with fluent condition set to 0 and disfluent condition set to 1.

```{r}
#| echo: true
#| tbl-cap: "First few rows from our dataset"
# Read the CSV file located in the "MS/data" directory and store it in a dataframe
fluency_data <- read.csv(here::here("MS/data/miko_data.csv")) %>%
  
  # Rename the columns for better readability
  rename(
    "Participant" = "ResponseID", # Rename "ResponseID" to "Participant"
    "Fluency" = "Condition",      # Rename "Condition" to "Fluency"
    "Time" = "name",              # Rename "name" to "Time"
    "Accuracy" = "value"          # Rename "value" to "Accuracy"
  ) %>%
  
  # Transform the data
  mutate(
    Accuracy = Accuracy *10 / 100, # Convert Accuracy values to proportions
    Fluency = ifelse(Fluency == 1, "Fluent", "Disflueny"), # rename levels 
    Fluency_dummy = ifelse(Fluency == "Fluent", 0, 1),  # Recode
    #Fluency: 1 becomes 0, others become 1 
    Fluency_dummy = as.factor(Fluency_dummy) # turn fluency cond to dummy code

  ) %>%
  
  filter(Time=="FreeCt1AVG") %>% # only choose first time point
  
  # Drop the column "X" and "time" from the dataframe
  select(-X, -Time)  %>%
  
  # move columns
  
  relocate(Accuracy, .after = last_col())
  

# Display the first few rows of the modified dataframe
head(fluency_data) %>%
  tt()
```

|               |                                  |
|---------------|----------------------------------|
| Column        | Key                              |
| Participant   | Participant ID number            |
| Fluency       | Fluent vs. Disfluent             |
| Fluency_dummy | Fluent: 0; Disfluent 1           |
| Accuracy      | Proportion recalled (idea units) |

: Columns and values from the dataset we are using

## OLS Regression

We first fit a regression model using the `lm`function to the data looking at final test accuracy (`Accuracy`) as a function of instructor fluency (`fluency_dummy`) Below is the code needed to fit the model in R.

```{r}
#| echo: true
#| warning: false
#| message: false

ols_model <- lm(Accuracy~Fluency_dummy, data=fluency_data)

```

```{r}
#| tbl-cap: "OLS  Regression Model coefficents"
#| tbl-cap-location: top

#| 
ols_model_new <- model_parameters(ols_model)

ols_model_new %>%
  print_html 

```

Focusing on output from our regression analysis in Table 3, we see the that there is a significant effect of Fluency, b = `r round(ols_model_new$Coefficient[2], 3)` , SE = `r round(ols_model_new$SE[2], 3)` , 95% CIs = `r round(ols_model_new$CI_low[2], 3)`,`r round(ols_model_new$CI_high[2], 3)`, p = `r round(ols_model_new$p[2], 3)`. This is exactly what @wilford2020 computed in their paper.

```{r}
#| fig-cap: "Proprotion Recalled on Final Test as a fucntion of Fluency"
#| echo: false

library(gghalves)

fluency_halves <- ggplot(fluency_data, aes(x = Fluency, y = Accuracy)) +
  geom_half_point(aes(color = Fluency), 
                  transformation = position_quasirandom(width = 0.1),
                  side = "l", size = 1, alpha = 0.5) +
  geom_half_boxplot(aes(fill = Fluency), side = "r") + 
  scale_y_continuous(labels = label_percent()) +
  scale_fill_viridis_d(option = "plasma", end = 0.8) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  guides(color = "none", fill = "none") +
  labs(x = "Fluency", y = "Pecentage Recalled on Final Test") +
  theme_lucid(base_size=14)

fluency_halves


```

## Beta regression approach

Using a traditional approach, instructor fluency impacts actual learning. Keep in mind the traditional approach assumes normality of residuals and homoscadacity. Does our model meet those assumptions? Using `easystats` [@easystats-2] and `check_model` function, we can easily assess this. As can be seen in Figure 5, there are some issues. Specifically, there is appears to violations of normality and homoscakdacity assumptions.

```{r}
#| label: fig.5
#| fig-cap: "Assumption Checks for `ols` model"
check_model(ols_model)
```

One solution is to run a beta regression model. Below we fit a beta regression using the `betareg` package[@betareg]. This a popular package for running frequentist beta regressions.

```{r}

# load betareg package
library(betareg)
```

```{r}
#| echo: true
#| error: true

# running beta model error
beta_model<-betareg(Accuracy~Fluency_dummy, data=fluency_data)
```

When you run the above model, an error will appear: `Error in betareg(Accuracy ~ Fluency_dummy, data = fluency_data) : invalid dependent variable, all observations must be in (0, 1)`. If your remember, the beta distribution can model responses in the interval \[0-1\], but not exactly 0 or 1. We need make sure there are no zeros and ones in our dataset.

Let's take a look at how many zeros and ones are in our dataset.

```{r}
fluency_data %>%
  filter(Accuracy==0 |Accuracy==1) %>%
  count(Accuracy) %>%
  tt()
```

In our dataset, we have 9 rows with accuracy of 0, and 1 row with an accuracy of exactly 1. To run a beta regression, we can employ a little hack. We can nudge our 0s towards .01 and 1s to .99 so they fall within the interval of \[0-1\].

```{r}
#| echo: true
#| 
# transfrom 0 to 0.1 and 1 to .99
data_beta <- fluency_data %>% 
    mutate(Accuracy = ifelse(Accuracy == 0, .01, ifelse(Accuracy == 1, .99, Accuracy)))

```

Let's fit the model again.

```{r}
#| echo: true
#| 
# fit beta model without 0s and 1s in our dataset 

beta_model<-betareg(Accuracy~Fluency_dummy, data=data_beta)

```

No errors this time! Now, Let's interpret the results of our beta regression. Table X includes the output of our beta regression model.

```{r}
#| tbl-cap: Model ouptput for our beta regression  model
#| tbl-cap-location: top

beta_model_results <- model_parameters(beta_model, component = "all")

beta_model_results %>%
  print_html()
```

### Model parameters

#### $\mu$ component

Looking at the model output in Table 4, there are two components we need to interpret when using beta regression. The first set of coefficients under fixed effects represents how factors influence the $\mu$ parameter, which is the mean of the beta distribution. These coefficients are interpreted on the scale of the logit link function, meaning they represent changes in the log-odds of the mean proportion. The intercept term represents the log odds of the mean on accuracy for the fluent instructor condition which we coded as the reference. Here being in the fluent condition translates to a log odds of `r round(beta_model_results$Coefficient[1], 3)`. The fluency coefficient represents the difference between the fluency and disfluency conditions. That is, watching a fluent instructor leads to higher recall than watching a disfluent instructor, b = `r round(beta_model_results$Coefficient[2], 3)` , SE = `r round(beta_model_results$SE[2], 3)` , 95% CIs = `r round(beta_model_results$CI_low[2], 3)`,`r round(beta_model_results$CI_high[2], 3)`\], p = `r round(beta_model_results$p[2], 3)`.

##### Predicted probabilities

While it is quite common to report the logits from a non-linear model, they are not very useful or intuitive. Instead we should discuss the effects of the predictor on the actual outcome of interest (in this case the 0-1 scale). The logit link allows us to transform back and forth between log-odds and probabilities. By using the inverse of the logit, we can easily transform coefficients to obtain proportions or percentages. In a simple case, we can do this manually, but when we have many moving pieces it can get quite complicated. Thankfully, there is a package called **`marginaleffects`** [@marginaleffects] that can help us extract the probabilities quite easily.[^2] To get the percentages for each of our categorical predictors we can use the function from the package called `avg_predictions`.

[^2]: `ggeffects` is another great package to extract marginal effects and plot [@ggeffects-2]

```{r}
#| echo: true
#load marginaleffects package
library(marginaleffects)
```

```{r}
#| echo: true
#| tbl-cap: Predicted probablities for fluency factor
#| tbl-cap-location: top

# get the predicted probablities for each level of fluency 
avg_predictions(beta_model, variables="Fluency_dummy") %>%
  tt()
```

Looking at table X, we see that both estimates are negative, which indicates that probability is below 50%. Looking at the predicted probablities confirms this. For the Fluency factor, we can interpret the estimate column in terms of proportions or percentages. That is, participants who watched the fluency video scored on average 36% on the final exam compared to 29% who watched the disfluency video.

##### Marginal effects

We can also examine changes in predicted probabilities for variations or small changes in an independent variable. This is called a marginal effect. There are different types of marginal effects, and various packages calculate them differently. Since we will be using the `marginaleffects` package for this tutorial, we will focus on the average marginal effect (AME), which is used by default in the `marginaleffects` package. The AME is the predicted change in the outcome for a very small change in the independent variable. In the `marginaleffects` package this invovle generating predictions for each row of the original data then averaging these predictions. One effect size measure we can calculate with categorical variables is the risk difference, which is the discrete difference between the average marginal effect of one condition or group and that of another condition or group. In the `marginaleffects` package, we can use the function `avg_comparisons` to obtain this metric. This function can also be used to get other popular effect size metrics, such as odds ratios and risk ratios.

```{r}
#| echo: true
#| tbl-cap: Margianl effects (risk difference) for Fluency factor
#| tbl-cap-location: top

# get risk difference 
beta_avg_comp<- avg_comparisons(beta_model, comparison= "difference") 

beta_avg_comp %>%
  tt()
```

Interpreting this output, In the fluency condition, participants who watched a fluent instructor scored 9% higher on the final recall test than participants who saw the disfluent instructor. This difference is reliably different, b= `r beta_avg_comp$estimate`, SE = `r beta_avg_comp$std.error`, 95 % CIs \[`r beta_avg_comp$conf.low`, `r beta_avg_comp$conf.high` \], p = `r beta_avg_comp$p.value`.

We can also get the odds ratio.

```{r}
#| echo: true
#| tbl-cap: Margianl effects (odds ratio) for Fluency factor
#| tbl-cap-location: top

# get odds ratios as an example
avg_comparisons(beta_model, comparison = "lnoravg",
    transform = "exp") %>%
  tt()

```

#### Plotting

We can easily plot the $\mu$ parmeter part of model using the `marginaleffects` package and the `plot_predictions` function.

```{r}
#| echo: true
#| fig-cap: Predicted probablities for Fluency factor

plot_predictions(beta_model, condition="Fluency_dummy")

```

#### Precision ($\phi$) component

The other component we need to pay attention to is the dispersion or precision parameter coefficients labeled as `phi` in Table 4. The $\phi$ parameter tells us how precise our estimate is. Specifically, $\phi$ in beta regression tells us about the variability of the response variable around its mean. Specifically, a higher dispersion parameter indicates a narrower distribution, reflecting less variability. Conversely, a lower dispersion parameter suggests a wider distribution, reflecting greater variability.

Understanding the dispersion parameter helps us gauge the precision of our predictions and the consistency of the response variable. In `beta_model` we only modeled the dispersion of the intercept. When $\phi$ is not specified, the intercept is modeled by default, as seen below.

```{r}
#| echo: true
#| 
beta_model<-betareg(Accuracy~Fluency_dummy, data=data_beta)
```

```{r}
#| echo: true
#| tbl-cap: Beta model output $\phi$ paramter
 

# get the precision paramter 
beta_model %>%
  model_parameters(component = "precision") %>%
  print_html()

```

The intercept under the precision heading is not that interesting. Let's model the dispersion of the `Fluency`factor. To do this we add a vertical bar to our `betareg` function which allows us to model the dispersion of any factor to the right of it. In the below model, `beta_model_dis`, we model the precision of the `Fluency` factor.

```{r}
#| echo: true

# add disp/percison for fluency 
beta_model_dis<-betareg(Accuracy~Fluency_dummy | Fluency_dummy, data=data_beta)
```

```{r}

#| tbl-cap: Beta model output with $\phi$ parameter for fluency factor


beta_model_dis %>%
  model_parameters(component = "all") %>%
  print_html()
```

Looking at the precision parameter coefficient, it is important to note that this estimate is on the logit scale, instead the estimates are logged. To interpret them on the original scale, we can exponent the log-transformed value—this transformation gets us back to our original scale. We get only the dispersion parameter by setting by setting the `component` argument to `precision` in `model_paramters`. We can also get the original value by including the `exponentiate = TRUE`.

```{r}
# get dis only from model
# exp to get original metric
beta_model_dis_results <- beta_model_dis %>% model_parameters(component = "precision")

beta_model_dis %>%
  model_parameters(component = "precision", exponentiate = TRUE) %>%
  print_html()
```

The $\phi$ intercept represents the precision of the fluent condition. The $\phi$ coefficient for `Fluency` represents the change in that precision for the Fluent instructors vs. Disfluent instructors. Here it is statistically significant, b = `r round(beta_model_dis_results$Coefficient[2], 3)` , SE = `r round(beta_model_dis_results$SE[2], 3)` , 95% CIs = `r round(beta_model_dis_results$CI_low[2], 3)`,`r round(beta_model_dis_results$CI_high[2], 3)`\], p = `r round(beta_model_dis_results$p[2], 3)`.

We have all the parameters to draw two different distributions of our outcome, split by fluency of the instructor. Let’s plot these two predicted distributions on top of the true underlying data and see how well they fit. In Figure 6, we can see how the distribution changes when we include a dispersion parameter for Fluency.

::: callout-note
Horrible fit. Why?
:::

```{r}
#| label: dispersion viz
#| echo: false

beta_model_dis_output  <- beta_model_dis %>% model_parameters(component = "all")

beta_mu_intercept <- beta_model_dis_output %>% as.data.frame() %>%
  filter(Component == "conditional", Parameter == "(Intercept)") %>% 
  pull(Coefficient)

beta_mu_fluency <- beta_model_dis_output  %>% 
   as.data.frame() %>%
   filter(Component == "conditional", Parameter == "Fluency_dummy1") %>% 
  pull(Coefficient)

beta_phi_intercept <- beta_model_dis_output  %>% 
  as.data.frame() %>%
   filter(Component == "precision", Parameter == "(Intercept)") %>% 
  pull(Coefficient)

# get phi for fluency
beta_phi_fluency <- beta_model_dis_output %>% 
  as.data.frame() %>%
   filter(Component == "precision", Parameter == "Fluency_dummy1") %>% 
  pull(Coefficient)


Fluency_same_prescion <- paste0("dprop(mean = plogis(", round(beta_mu_intercept, 2),
                         "), size = exp(", round(beta_phi_intercept, 2), "))")


Fluency_different_precison <- paste0("dprop(mean = plogis(", round(beta_mu_intercept, 2),
                      " + ", round(beta_mu_fluency, 2), 
                      "), size = exp(", round(beta_phi_intercept, 2),
                      " + ", round(beta_phi_fluency, 2), "))")

ggplot(data = tibble(x = 0:1), aes(x = x)) +
  geom_density(data = data_beta, 
               aes(x = Accuracy, fill = Fluency_dummy), alpha = 0.5, color = NA) +
  stat_function(fun = dprop, size = 1,
                args = list(size = exp(beta_phi_intercept), 
                            mean = plogis(beta_mu_intercept)),
                aes(color = Fluency_same_prescion)) +
  stat_function(fun = dprop, size = 1,
                args = list(size = exp(beta_phi_intercept + beta_phi_fluency), 
                            mean = plogis(beta_mu_intercept + beta_mu_fluency)),
                aes(color =Fluency_different_precison )) + 
 scale_fill_viridis_d(option = "mako", end = 0.8, name = "Fluency",
                       guide = guide_legend(ncol = 1, order = 1)) +
  scale_color_viridis_d(option = "mako", end = 0.8, direction = -1, name = "",
                        guide = guide_legend(reverse = TRUE, ncol = 1, order = 2)) + 
    scale_x_continuous(labels = label_percent()) +

  labs(x = "Accuracy Final Recall") +
  theme(legend.position = "bottom")


```

```{r}
#| label: no dispersion viz
#| echo: false


# get phi for fluency
beta_phi <- beta_model_results %>% 
  as.data.frame() %>%
   filter(Component == "precision", Parameter == "(phi)") %>% 
  pull(Coefficient)


Fluency_same_prescion <- paste0("dprop(mean = plogis(", round(beta_mu_intercept, 2),
                         "), size = ", round(beta_phi, 2), "))")


Fluency_different_precison <- paste0("dprop(mean = plogis(", round(beta_mu_intercept, 2),
                      " + ", round(beta_mu_fluency, 2), 
                      "), size = ", round(beta_phi, 2), ")")

ggplot(data = tibble(x = 0:1), aes(x = x)) +
  geom_density(data = data_beta, 
               aes(x = Accuracy, fill = Fluency_dummy), alpha = 0.5, color = NA) +
  stat_function(fun = dprop, size = 1,
                args = list(size = beta_phi, 
                            mean = plogis(beta_mu_intercept)),
                aes(color = Fluency_same_prescion)) +
  stat_function(fun = dprop, size = 1,
                args = list(size = beta_phi, 
                            mean = plogis(beta_mu_intercept + beta_mu_fluency)),
                aes(color =Fluency_different_precison )) + 
 scale_fill_viridis_d(option = "mako", end = 0.8, name = "Fluency",
                       guide = guide_legend(ncol = 1, order = 1)) +
  scale_color_viridis_d(option = "mako", end = 0.8, direction = -1, name = "",
                        guide = guide_legend(reverse = TRUE, ncol = 1, order = 2)) + 
    scale_x_continuous(labels = label_percent()) +

  labs(x = "Accuracy Final Recall") +
  theme(legend.position = "bottom")

```

Note how the whole distribution changes when we include precision of `Fluency`. The data doesn't fit the underlying distribution very well. Despite this, this makes clear the importance of including a precision parameter. A critical assumption of the GLM is homoscedasticity, which means constant variance of the errors. Here we see one of the benefits of a beta regression model. We can include a dispersion parameter for Fluency. Properly accounting for dispersion is crucial because it impacts the precision of our mean estimates and, consequently, the significance of our coefficients. The inclusion of dispersion in the our model changed the statistical significance of the $\mu$ coefficient. This suggests that failing to account for the dispersion of the variables might lead to biased estimates. This highlights the potential utility of an approach like beta regression over a traditional GLM (regression or ANOVA approach), as beta regression can explicitly model dispersion and address issues of heteroscedasticity.

We wont always need to include dispersion parameters for each of our variables. it is preferable to conduct a very simple likelihood ratio test (LRT) to examine if we need to include dispersion into our model. To test this we use the `test_likelihoodratio` from the `easystats` ecosystem [@easystats].

```{r}
#| echo: true
#| 

beta_model <- betareg(Accuracy~Fluency_dummy , data=data_beta)

beta_model_dis<-betareg(Accuracy~Fluency_dummy | Fluency_dummy, data=data_beta)


test_likelihoodratio(beta_model, beta_model_dis) %>%
  tt()

```

Here we want to model the precision for fluency.

## Bayesian implementation of beta regression

We can also fit beta regression models under a Bayesian framework. Adopting a Bayesian framework often provides more flexibility and allows us to quantity uncertainty around our estimates which makes it more powerful than the frequentist alternative. For the purposes of this tutorial, we will not be getting into the minutae of Bayesian data analysis (i.e., setting informative priors, MCMC sampling, etc,). For a more in-depth look into Bayesian data analysis I refer the reader to XXX.

For the following analyses we will be using default priors provided by brms. This will get us something tantamount to a frequent analysis most of the readers are used to seeing.

To fit our Bayesian models, we will be using a Bayesian package called `brms` [@brms-2] . b`rms` is a powerful and flexible Bayesian regression modeling package that offers built in support for the beta distribution and some of the alternatives we discuss in this tutorial.

We can recreate the beta model from `betareg` in `brms` easily. Instead of using the `|` operator to specify different parameters, we model each parameter independently. Recall we are fitting two parameters— $\mu$ and $\phi$. We can easily do this by using the `bf` function from `brms`. `bf()` facilitates the specification of several sub-models within the same formula call. We fit two formulas, one for $\mu$ and one for $\phi$ and store it in `model_beta_bayes`. Here we allow `Fluency_dummy` ro

```{r}
#| echo: true
#| 
#load brms and cmdstanr
library(brms)
library(cmdstanr) # Use the cmdstanr backend for Stan because it's faster and more modern than
  # the default rstan You need to install the cmdstanr package first
  # (https://mc-stan.org/cmdstanr/) and then run cmdstanr::install_cmdstan() to
  # install cmdstan on your computer.
options(marginaleffects_posterior_center = mean)
# get mean instead of mediation in marginaleffects
```

```{r}
# fit model with mu and phi
model_beta_bayes <- bf(Accuracy  ~ Fluency_dummy, # mu 
     phi ~ Fluency_dummy) # phi 
```

We can then pass `model_beta_bayes` to the `brm` fucntion and set the model family to the beta distribution, which is native to our model using the `brm` function. We also set a bunch of arguments to speed up the fitting of the models.

```{r}
#| label: van beta model
#| results: hide

# run model 

beta_brms <- brm(model_beta_bayes, 
    data = data_beta,
  family = Beta(),
  chains = 4, iter = 2000, warmup = 1000,
  cores = 4, seed = 1234, 
  backend = "cmdstanr",
  file = "model_beta_bayes_1"  # Save this so it doesn't have to always rerun
)
```

If you recall we fit the model with default priors. You can access the priors used by using the `prior_summary` function

```{r}
#| label: set prior for beta model

prior_summary(beta_brms) %>%
  tt()

```

```{r}
#| tbl-cap: Parameter estimates for beta regression using `brms`

model_parameters(beta_brms, "mean")%>%
  tt()

```

Parameter estimates for `brms` fitted beta regression model are labeled in Table X.[^3] To make the output more readable, each model parameter is labeled with a prefix, except for the $\mu$ parameter, which takes same name as the variables in your model. Comparing results with \``` betareg` ``in `model_beta`, our results are very similar. Additionally, the parameters can be interpreted in a similar manner and we can use `marginaleffects` to extract marginal effects and risk difference. [^4]In contrast to frequentist models we can interpret this in probabilistic terms. There is no *p*-values to interpret in Bayesian models. What is sometimes done to judge statistical significance is examine the 95% credible interval to see if it includes 0--if it does not then the effect can be said to be significant. The results are basically the same that we found with \``` betareg` ``in `model_beta`.

[^3]: We have chain diagnostics included like Rhat and ESS which indicates how the MCMC sampling performed. For more information check out Gelman et al., 2013; Kruschke, 2014; McElreath, 2020)

[^4]: It is important to note that these transformations should be applied to draws from the posterior distribution. `Marginaleffects` does this under the hood, but other packages might not.

### Posterior predictive check

We can look at the fit of our model and see how well it fits our data by using the `pp_check` function. Looking at Figure X, The x-axis represents the possible range of outcome values and the y-axis represents the density of each outcome value. Ideally, the predictive draws (the lighly blue lines) should show reasonable resemblance with the observed data (dark blue line). We see it does a pretty good job capturing the data.

```{r}
#| label: ppcheck beta model

pp_check(beta_brms,ndraws = 100)
```

### Predicted probabilities

```{r}
beta_brms %>% 
  avg_predictions(variables = "Fluency_dummy") %>%
  tt()
```

### Marginal effects

```{r}
# risk difference
beta_brms  %>% 
  avg_comparisons(variables = "Fluency_dummy") %>%
  tt()
```

The risk difference in predicted outcomes of 0.05, which is roughly what we found in earlier models. We also have a 95% credible interval for the difference. The 95% credible interval does not include zero so we can it is statistically significant.

### Plotting

Similar to our frequentist model, we can use `plot_predictions` to plot our model on the original scale.

```{r}
#| echo: true
#| fig-cap: Predicted probablities for Fluency factor with 95% Credible Intervals

plot_predictions(beta_brms, condition="Fluency_dummy") + theme_lucid(base_size = 14) +
scale_y_continuous(labels = label_percent())

```

## Zero-inflated beta (ZIB) regression

A limitation of the beta regression model is it can can only model values between 0 and 1, but not 0 or 1. In our dataset we have 9 rows with `Accuracy` equal to zero.

To use the beta distribution we nudged our zeros to 0.01--which is never a good idea in practice. In our case it might be important to model this, as fluency of instructor might be an important factor in predicting the zeros in our model. Luckily, there is a model called the zero-inflated beta (ZIB) model that takes into account the structural 0s in our data. We’ll still model the 𝜇 and 𝜙 (or mean and precision) of the beta distribution, but now we’ll also add one new special parameter: 𝛼. With zero-inflated regression, we’re actually modelling a mixture of data-generating process. The $\alpha$ parameter uses a logistic regression to model whether the data is 0 or not. Below we fit a model called `beta_model_0` using the `glmmTMB` package. The `betareg` model cannot model zero-inflated data. In the `glmmTMB` function, we can model the zero inflation by including an argument called `ziformula`. This allows us to model the new parameter $\alpha$. Let's fit a model where there is a zero-inflated component for `Fluency`.

```{r}

# keep 0 but transform 1 to .99
data_beta_0 <- fluency_data %>% 
    mutate(Accuracy = ifelse(Accuracy == 1, .99, Accuracy))

```

```{r}
#| label: glmmTMB zib
#| echo: true
beta_model_0<-glmmTMB(Accuracy~Fluency_dummy, disp=~Fluency_dummy,  ziformula = ~ Fluency_dummy, data=data_beta_0, family=beta_family(link="logit"))
```

### Model parameters

```{r}
#| echo: true

model_zi <- model_parameters(beta_model_0) 

model_zi %>% 
 print_html()
```

As before, the 𝜇 parameter coefficients under the Fixed Effects header (i.e., `Intercept`, `Fluency_dummy)` are on the logit scale; while 𝜙 parameter coefficients under the Dispersion header are on the log scale. In addition, the paramters under the zero-inflated header are on the logit scale. Looking at the output, the $\mu$ part of the model, there are no significant effects. However, for the zero-inflated part of the model, the`Fluency` predictor is significant.

### Predicted probabilities and marginal effects

Because is on the logit scale we can back-transform it to the probability scale. To do this easily, we can use the `avg_predictions` function from `marginaleffects` package. Because we are interested in the zero-inflated part of the model we set the `type` argument to `zprob`.

```{r}
#| echo: true
#| 
beta_model_0 %>%
  marginaleffects::avg_predictions(by = "Fluency_dummy", type="zprob") %>%
  tt()

```

The estimates provided in Table X are percentages of zeros in the model. There are 2% 0s in the fluent condition and 15% 0s in the disfluent condition. If we calucalted the proprotion of zeros in the actual dataset.

We can also get the average marginal effect of Fluency like we did before:

```{r}
#| echo: true
#| 
 beta_model_0 %>%
  marginaleffects::avg_comparisons(variables = "Fluency_dummy", type="zprob", comparison = "difference")

```

Interpreting the estimate, the disfluency condition has a higher percentage of zeros (50%) than the fluent condition (2%). Looking at the risk difference, seeing lecture videos with a fluent instructor reduces the proportion of zeros by about 13%. Here we have evidence that participants were more likely to do more poorly after watching a disflueny lecture than a fluent lecture.

As a word of caution, the `marginaleffects` package has some issues with fitting models from `glmmTMB` and it has been recommended not to use it. Thus, we will not highlight it here. Please see this issue for more information: <https://github.com/vincentarelbundock/marginaleffects/issues/1064>

## Bayesian Implementation of zero-inflated beta

Luckily, we can fit a zero-inflated model using brms and use the `marginaleffects` package to make inferences about our parameters of interest. Similar to our beta model we fit in `brms` we will use the `bf()` function to fit several models. We fit our $\mu$ and $\phi$ parameters as well as our zero-inflated parameter ($\alpha$; here labeled as `zi`). In `brms` we can use the zero_inflated_beta family argument.

```{r}
#| label: zib brms model
#| echo: true
#| 

zib_model <-   bf(
    Accuracy ~ Fluency_dummy,  # The mean of the 0-1 values, or mu
    phi ~ Fluency_dummy,  # The precision of the 0-1 values, or phi
    zi ~ Fluency_dummy,  # The zero-or-one-inflated part, or alpha
  family = zero_inflated_beta()
)

```

We pass `zib_model` to the `brm` function.

```{r}
#| label: fit zib model
#| results: hide
#| echo: true
fit_zi <- brm(
  formula = zib_model,
  data = data_beta_0, 
  cores = 4,
  iter = 2000, 
  warmup = 1000, 
  seed = 1234, 
  backend = "cmdstanr",
  file = "model_beta_bayes_zib"
)

```

```{r}
#| tbl-cap-location: "top"
#| tbl-cap: Paramter estimates for zero-inflated beta model
#| 
model_parameters(fit_zi, "mean")%>%
  tt()
```

To get the zero-inflated part of the model we can set `dpar` argument to `zi` and get risk difference between the conditions. With a Bayesian implementation we get 95% credible intervals and we see the difference is significant.

We can easily plot the zero-inflated part of the model using the `plot_predictions` function

```{r}
#| label: plot brms zi model
#| fig-cap: "Predicted zero-inflated probablities of fluency on final test accuracy"
#| warning: false
#| message: false
#

zi_plot <- plot_predictions(fit_zi, by = c("Fluency_dummy"), dpar="zi")


```

## Zero-one inflated beta (ZOIB)

Sometimes it is theoretically useful to model zeros and ones in our data, For example, this could be important in visual analog scale data where there might be a prevalence of responses at the bounds [@vuorre2019] or free-list task where individuals provide open responses to some question or topic that are then recoded to fall between 0-1 [@bendixen2023]. In our case, we only have one participant with a perfect score, but in other datasets using accuracy as a DV this might not be the case.

The ZIB model works well if you have 0s in your data, but not 1s. In our data, we have exactly one value equal to 1. While probably not significant to alter our findings, we can model ones with a special type of model called the zero-one inflated beta (ZOIB) model. Unfortunately, there is no frequentist implemnetation of the ZOIB model. Luckily, we can fit a Bayesian implementation of the ZOIB model in `brms`. In this model, we fit four parameters or sub-models. We fit seperate models for the mean (\$\\mu\$) and the precision (\$\\phi\$) of the beta distribution; a zero-one inflation parameter (i.e. the probability that an observation is either 0 or 1; $\alpha$ ); and a ‘conditional one inflation’ parameter (i.e. the probability that, given an observation is 0 or 1, the observation is 1; $\gamma$). This specification captures the entire range of possible values while still being constrained between zero and one.

We use the `bf` function again to fit models for our four parameters. We use the native zero_one_inflated_beta family to fit our model.

```{r}

zoib_model <-   bf(
    Accuracy ~ Fluency_dummy,  # The mean of the 0-1 values, or mu
    phi ~ Fluency_dummy,  # The precision of the 0-1 values, or phi
    zoi ~ Fluency_dummy,  # The zero-or-one-inflated part, or alpha
    coi ~ Fluency_dummy,   # The one-inflated part, conditional on the 0s, or gamma
  family = zero_one_inflated_beta()
)


```

We then pass the `zoib_model` to our `brm` function.

```{r}
#| label: fit zoib model
#| results: hide
#| message: false
#| 

fit <- brm(
  formula = zoib_model,
  data = fluency_data, 
  chains = 4, iter = 2000, warmup = 1000,
  cores = 4, seed = 1234, 
  backend = "cmdstanr",
  file = "model_beta_zoib_1" 
)
```

```{r}
zoib_model <- parameters::model_parameters(fit, "mean")

zoib_model %>%
  print_html()
```

The output for the model is pretty lengthy—we are estimating four parameters each with their own independent models. All the coefficients are on the logit scale, except 𝜙, which is on the log scale.Thankfully drawing inferences for all these different parameters, plotting their distributions, and estimating their average marginal effects looks exactly the same as what we did earlier with the zero-inflated model—all the **brms**, and **marginaleffects** functions we used work the same.

### Plotting ZOIB

```{r}

```

## Ordered beta regression

Looking at the output from the ZOIB model in Table X, we can see how running a ZOIB model can become vastly complex and computational intensive with larger models. A special version of the ZOIB was recently developed called ordered beta regression [@kubinec2022]. The ordered beta regression model allows for the analysis of continuous data with upper and lower boundaries (e.g., 0 and 1). The ordered beta regression model is a hybrid model that combines a beta model with ordinal logistic regression model. An in-depth explanation of ordinal regression is beyond the scope of this tutorial (but see [@Fullerton2023; @bürkner2019]). At a basic level, ordinal regression models are useful for outcome variables that are categorical in nature and have some inherent ordering (e.g., Likert scale items or scales). To preserve this ordering, ordinal models relie on the cumulative distribution. Within an ordinal regression model, going from one level or category to another is modeled with a single set of covariates that predicts cutpoints between each category. That is, each coefficient shows the effect of moving from one option to a higher option with *k*-1 cutpoint parameters showing the boundaries or thresholds between the probabilities of these categories. Since there’s only one underlying process, there’s only one set of coefficients to work with (proportional odds assumption). In an ordered beta regression, three ordered categories are modeled: (1) exactly zero, (2) somewhere between zero and one, and (3) exactly one. In an ordered beta regression, (1) and (2) are modeled with cumlative logits, where one cutpoint (`cutzero`) is the the boundary between *Exactly 0* and *Between 0 and 1 and the other cutpoint* (`cutone`) is the boundary between *Between 0 and 1* and *Exactly 1. Somewhere between 0-1 (3)* is modeled as a beta regression with parameters reflecting the mean response on the logit scale. The ordered beta regression model has shown to be more efficient than some of the methods discussed herein and deserves special mention.

## Frequentist implementation

We can run an ordered beta regression using the `glmmTMB` function and changing the family argument to `ordbeta`.

```{r}
#| label: run ordered beta regression
#| 
ord_fit <- glmmTMB(formula=Accuracy ~ Fluency_dummy  + (1|Participant),
         data=fluency_data,
         family = ordbeta)
```

```{r}

ord_fit %>%
  model_parameters() %>%
  print_html()
```

If we take a look at the summary output, we can interpret the values similar to a beta regression. Remember these values are on the logit scale so we can take the inverse and get probabilities like we have done above. We can easily do this with the `marginaleffects` package. The only significant effect is Time:

```{r}

# get the predicted probablities for each level of fluency 
avg_predictions(ord_fit, variables="Fluency_dummy") %>%
  print_html()

```

We can get the risk difference between the two time points:

```{r}

# get risk difference 
avg_comparisons(ord_fit,variables="Fluency_dummy",  comparison= "difference") %>%
  tt()

```

## Bayesian implementation

To fit an ordered beta regression in a Bayesian context you can use `ordbetareg` [@ordbetareg] package. This uses `brms` on the front-end and is straightforward to run.

```{r}

#| label: load ordbetareg

library(ordbetareg)

```

```{r}
#| label: brms ordered beta model
#| results: hide
#| echo: true


# use ordbetareg to fit model
ord_fit_brms <- ordbetareg(Accuracy ~ Fluency_dummy,
                      data=fluency_data,
                      chains=4,
                      iter=2000,
                      backend="cmdstanr", 
                      file = "model_beta_ordbeta")


```

```{r}
#| tbl-cap: Model parameters for ordered beta model
#| tbl-cap-location: "top"
#| 
ord_fit_brms %>%
  model_parameters() %>%
  tt()

```

### Ordered beta model fit

The best way to visualize model fit is to plot the full predictive distribution relative to the original outcome. Because ordered beta regression is a mixed discrete/continuous model, a separate plotting function, `pp_check_ordbetareg`, is included in the `ordbetareg` package that accurately handles the unique features of this distribution. This function returns a list with two plots, `discrete` and `continuous`, which can either be printed and plotted or further modified as `ggplot2` objects.

```{r}
#| label: Model fits
#| echo: true

plots <- pp_check_ordbeta(ord_fit_brms,
                          ndraws=100,
                          outcome_label="Final Test Accuracy")


```

The discrete plot, which is a bar graph, shows that the posterior distribution accurately captures the number of different types of responses (discrete or continuous) in the data.

```{r}

plots$discrete
```

For the continuous plot, shown as a density plot with one line per posterior draw, the model does a very good job at capturing the distribution.

```{r}
plots$continuous
```

It is clear from the posterior distribution plot that the ordered beta model fits the data well.

# Discussion and Conclusion

The use of beta regression in psychology, and the social sciences in general, is rare. With this tutorial we hope to change this. Beta regression models are an attractive alternative to models subsumed under the GLM, which imposes the unrealistic assumptions of normality, homoscadacidty, requires the data to be unbounded. There is actually a diverse array of different models that can be used depending on your outcome of interest. Throughout this tutorial our main aim was to help guide researchers in running a beta regression with real data by providing R code and providng an in-depth walk through on how to interepret model parameters for each model, as well as how to extract predicted probablities and marginal effects from each model, and finally visualize the results from each model. The empirical example in this study, accuracy on the final test, is ideally suited for a beta model. There are a whole host of outcomes used in psychology that would be ideal for beta regression and its alternatives. Researchers now have some tools to analyze their data properly.

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

# Appendix

```{r}

library(grateful)

```

# Title for Appendix
